---
title: "P lividus larvae light levels"
author: "John Kirwan"
date: "3/19/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
 
```{r include=FALSE}
library('dplyr')
library('tidyr')
library('ggplot2')
```


> Male and female individuals aggregate for spawning and simultaneously release their gametes (Cherbonnier, 1954). In Port-Cros Bay, France, spawning aggregations of 10 to 20 individuals occur at dusk on prominent stones or at the top of leaves of *P. oceanica*.

> In the northwestern Mediterranean, larvae are present in the plankton year-round. One or two peaks occur, the first in May to June, the second, if present, from September to November (Fenaux, 1968; Fenaux and Pedrotti, 1988; Pedrotti and Fenaux, 1992; Pedrotti, 1993). The duration of the planktonic life of larvae of P. lividus has been estimated to be 23 to 29 days in situ (Pedrotti, 1993; but see Lozano et al., 1995), but it can drop to 14 to 19 days in vitro with non-limiting food (George et al., 1989; Fenaux et al., 1992).

Need to determine the appropriate light level for sea urchin larvae during the night and day phase of their diel cycle. According to Maurizio, the larvae are probably found at 100 m depth during the day and close to the surface at night. Need to calculate the photon irradiance in the appropriate wavelength range and provide similar light to the animals using aquarium lights. 

Yet, the gulf of Naples is not 100 m deep - where are larvae likely to reside during the day? Moreover, if light is the sole guide - the light level of water will vary dramatically. 

MC is the Gulf of Naples, i01 is the Ionian sea, Nb4 is the Northern Mediterranea sea, and alt is the central Tyrrhenian sea. EMTEC is also the Ionian sea. 

LTER-MC: Marechiara sampling point.
> The sampling site LTER-MC is located 2 miles offshore the city of Naples, on the 75 m isobath, at the border between the coastal eutrophic system influenced by land runoff and the offshore oligotrophic waters with characteristics of the southern Tyrrhenian Sea.


#### Climatological irradiance data from Gulf of Naples

The following data come from Hydrolight, courtesy of Maurizio Ribera.

> Units are microEinstein m^-2 s^-1. E_d is downwelling (cosine) E0_ is scalar, both in absolute value at given depth. R and F are for Raman scattering a Fluorescence 0 not accounted for 1 accounted for (I am sending those with 1,1). The incident irradiance is the cloud free irradiance for that date at the site.

The values are **per nm in the binwidth range**

> The incident irradiance is the cloud free irradiance for that date at the site. However E_d(0) is the E_d(0^-). Since the profiles are all at noon you can assume that E_d(0^-) =0.98 E_d(0^+) the incident irradiance at sea surface.

The directed irradiance given as 0m is actually below the surface and should be 98% of the value just above the surface. 

The values include both Raman scattering and fluorescence. The incident irradiance is the cloud free irradiance for that date at the site at noon.

E<sub>d</sub> is *directed irradiance*. E<sub>0</sub> is *scalar irradiance*. 

> The technology of measurement of scalar irradiance is essentially the same as that of irradiance with the exception of the collector. Since this must respond equally to light from all directions, it has to be spherical. 

> Once you have absolutely calibrated your system, irradiance measurements are straightforward. Remember from chapter 2 that two kinds of irradiance are the most useful, vector and scalar. Vector irradiance measures light hitting a surface from all angles in one hemisphere, but weights the amount of the light from each direction by the cosine of its angle from the perpendicular to the surface. Scalar irradiance measures all the light intersecting a surface from the whole sphere, giving all directions equal weight. Both properties are measured using special sampling optics.

> Scalar irradiance does not suffer from the orientation dependence of vector irradiance. Scalar irradiance is given by the number of photons that pass through the surface of a small sphere over a given period of time. It has the same units as vector irradiance, but the directions of the photons do not matter. So, for any location in space there is only one scalar irradiance, while vector irradiance depends on the orientation of the detector.

For these purposes, scalar irradiance makes more sense.

Those who study phytoplankton use *scalar irradiance*. In studies of photosynthesis, one *Einstein* is a mole of photons.

The wavelength bins are 5 nm wide and the stated value is the midpoint of the bin, i.e. the minimum value of 352.5 refers to the 350 - 355 nm bin, whereas the maximum value 862.5 refers to the bin from 860 - 865 nm.  The values are nonetheless per nm. 

We bring in the data. I had interpolated the irradiance at 100 m by taking the decrement from 50 m to 75 m at each wv bin (25-30%) and applying this to the 25 m below. 

> It is definitely reasonable to obtain deeper value using the average $K(\lambda)$ over, say, 50 to 60 m. 
>
> -- <cite>Maurizio di Montesanto</cite>

Therefore, the interpolation is over the preceding 50 m, the proportion of light remaining being squared to get the change over the subsequent 25 m - the two results being very similar.

```{r message=FALSE, warning=FALSE}
#seq1 <- seq(from=0,to=75,by=5)
# E0_df <- readr::read_delim('MC_E0_R1F1.txt',delim=',')#,# col_names = c('wv',seq1),skip=1)
wide_sea <- readr::read_delim('light_level.csv',delim=';',col_types = "cfffn")

# colnames(E0_df)[1] <- 'wv' 
# #E0_df$`100` <- ( E0_df$`75` / E0_df$`50` ) * E0_df$`75`  # 25 m change at depth
# E0_df$`100` <- sqrt( E0_df$`75`/E0_df$`25`) * E0_df$`75` # 25 m change at depth
# E0_df$`125` <- ( E0_df$`75` / E0_df$`50` ) * E0_df$`100` # 25 m change at depth
# E0_df$`150` <- ( E0_df$`75` / E0_df$`50` ) * E0_df$`125` # 25 m change at depth
# head(E0_df)
```
 
Add a column with irradiance in quanta arther than umols thereof. 
 
```{r}
wide_sea |> tidyr::pivot_longer(cols = "0":"200",names_to = "depth",
                                values_to = "umol") |> 
  drop_na(umol) -> water_cols
water_cols$depth <- as.numeric(water_cols$depth)

water_cols |>
  mutate(quanta = (umol/1e6) * 6.02214076e23 ) -> water_cols
```


The data include measurements at E0 and Ed. They are similar, with the E0 slightly higher, so we will remove the Ed for simplicity.

```{r}
water_cols <- droplevels( water_cols[water_cols$E == "E0",])
```

  

What percentage of light remains compared to the surface?
 

```{r}
# propJoe <- function(umol,depth) {
#   prop <- 100*umol[depth==10]/umol[depth==0]
#   return(prop)
# }

water_cols |> group_by(sea,E,wv) |>
  filter(between(wv,300,700)) |>
  summarise(`10` = 100*umol[depth==10]/umol[depth==0],
            `20` = 100*umol[depth==20]/umol[depth==0],
            #`150` = 100*umol[depth==100]/umol[depth==0],
            #`200` = 100*umol[depth==200]/umol[depth==0] ,
            ) |>
    pivot_longer(cols= num_range("",0:200),names_to="depth",
                 values_to = "proportion" ) -> shallow
shallow$depth <- as.numeric(shallow$depth)

shallow |> 
  ggplot(aes(y=proportion,x=wv,color=sea)) + 
  facet_wrap(vars(depth),scales = "free_y") +
  geom_line(aes(linetype=E)) + 
  ggtitle('Percentage photons
          remaining at depths by wavelength') + theme_classic() +
  ylab('Surface photons remaining (%)') 
```

Where the MC (Naples) data cuts off is below:


```{r}
# propJoe <- function(umol,depth) {
#   prop <- 100*umol[depth==10]/umol[depth==0]
#   return(prop)
# }

water_cols |> group_by(sea,E,wv) |>
  filter(between(wv,300,700)) |>
  summarise(`100` = 100*umol[depth==100]/umol[depth==0],
            `200` = 100*umol[depth==200]/umol[depth==0] ,
            ) |>
    pivot_longer(cols= num_range("",0:200),names_to="depth", values_to = "proportion" ) -> deep
deep$depth <- as.numeric(deep$depth)

deep |> 
  ggplot(aes(y=proportion,x=wv,color=sea)) + 
  facet_wrap(vars(depth),scales = "free_y") +
  geom_line(aes(linetype=E)) + 
  ggtitle('Percentage photons remaining at depths by wavelength') + theme_classic() + ylab('Surface photons remaining (%)') 

# E0_df |>
#   filter(between(wv,300,700)) |>
#   mutate(light_left = 100*`100` / `0`) |>
#   ggplot(aes(y=light_left,x=wv)) + geom_line() + 
#   ggtitle('Percentage light remaining at 100 m depth from 300-700 nm') + theme_classic()
```

About 1% in the blue range - less elsewhere. The North Mediterranean water is turbid. 

We gather these data at various depths into a single *depth* variable. The quanta is the irradiance (converted to moles) by Avogrado's number. We multiply it by 5 (the binwidth),as values are per nm in the specfied range.

```{r}
# E0_df |> tidyr::gather(key="depth",value="irrad",-'wv') |> 
#   group_by(depth,wv) |>
#   mutate(irrad_SD = round(sd(irrad),2),
#          quanta   = (irrad / 1e6) * 6.02214076e23
#          ) -> scalar_irrad_df
# scalar_irrad_df$depth <- as.numeric(scalar_irrad_df$depth)
# head(scalar_irrad_df)
```

Plot the quanta for shallower water.

```{r}
water_cols |> group_by(E,sea) |>
  filter(E=='E0' & depth %in% c(0,10,20) & between(wv,350,650)) |>
 # filter(depth[depth%%2==0]) |>
  ggplot(aes(x=wv,y=quanta,color=-depth)) + geom_point() +
  theme_classic() + 
  scale_y_log10() + facet_wrap(vars(sea))#,scales = "free_y")
```

 Plot the quanta for deeper water.
 
```{r}
water_cols |> group_by(E,sea) |>
  filter(E=='E0' & depth %in% c(75,100,200) & between(wv,350,650)) |>
 # filter(depth[depth%%2==0]) |>
  ggplot(aes(x=wv,y=quanta,color=-depth)) + geom_point() +
  theme_classic() + 
  scale_y_log10() + facet_wrap(vars(sea))#,scales = "free_y")
```
 
This is approximately in keeping with S. Johnsen's Optics of life Absorption chapter, which gives similar values for light at 500 nm (using irradiance with units masured in sq cm.)
 
The bay of Naples is similar to the central Tyrrhenian and Ionian. 
 
 
```{r warning=FALSE}
# scalar_irrad_df |> group_by(depth) |>
#   filter(depth %in% c(0,5,10,25,50,75,100)) |>
#  # filter(depth[depth%%2==0]) |>
#   ggplot(aes(x=wv,y=irrad,color=-depth)) + geom_point() + theme_classic() #+
#   #scale_y_log10()
```
 
Focus on the deepest three conditions. 
 
```{r warning=FALSE}
# scalar_irrad_df |> group_by(depth) |>
#   filter(between(wv,350,700)) |>
#   filter(depth %in% c(50,75,100)) |>
#  # filter(depth[depth%%2==0]) |>
#   ggplot(aes(x=wv,y=irrad,color=-depth)) + geom_point() + theme_classic() #+
#   #scale_y_log10()
```
  
Get the relative decrement from 50 m to 75 m depth and apply this to 75 m depth to interpolate 100 m depth.
 
```{r warning=FALSE}
# scalar_irrad_df |> group_by(depth) |>
#   filter(wv < 700) |>
#   filter(depth %in% 100) |>
#  # filter(depth[depth%%2==0]) |>
#   ggplot(aes(x=wv,y=irrad)) + geom_point() + 
#   geom_col(width=5,alpha=.5,fill="lightblue",col="yellow") +
#  # geom_smooth(method = "loess", span=0.1) + 
#   theme_classic()
```
 
There is actually still some power below 400 nm (into near UV).
 
```{r}
# print(paste0('Percentage of photons below 400nm at 100 m depth: ',
#   signif(100 * sum(scalar_irrad_df$quanta[
#     scalar_irrad_df$wv < 400 & scalar_irrad_df$depth==100]) / 
#     sum(scalar_irrad_df$quanta[scalar_irrad_df$depth==0 & scalar_irrad_df$wv < 400]),
#   2),'%'))
```
 
A non-negligible amount of UV light remains at 100m. We will ignore these photons in finding a scalar value of irradiance that we can match with the PAR-sensitive (400-700 nm) irradiance meter at Stazione Zoologica Anton Dohrn. 
 
### Sum values
 
How many moles of photons (400 - 700) are present in the light at various depths?
 
 
```{r}
water_cols |>
  group_by(sea,depth) |>
  filter(E=='E0', between(wv,400,700)) |>
  summarise(irradiance = signif(sum(umol*5),4)) -> tot_phot_irrad

head(tot_phot_irrad)
```

 
```{r}
tot_phot_irrad |>
  ggplot(aes(x=depth,y=irradiance,colour=sea)) + geom_point() +
  geom_line() + scale_y_log10()  +
  theme_classic() + theme(legend.position = c(.15,.2))
```

The downwelling light at 5 m in Naples is:

```{r}
print(paste(
  tot_phot_irrad$irradiance[tot_phot_irrad$depth==5 & 
                            tot_phot_irrad$sea=="Naples"],
  'umol m^-2 s^-1.'
))
```
 
The downwelling light at 10 m in Naples is:

```{r}
print(paste(
  tot_phot_irrad$irradiance[tot_phot_irrad$depth==10 & 
                            tot_phot_irrad$sea=="Naples"],
  'umol m^-2 s^-1.'
))
```
At 75 m:
  
```{r}
print(paste(
  tot_phot_irrad$irradiance[tot_phot_irrad$depth==75 & 
                            tot_phot_irrad$sea=="Naples"],
  'umol m^-2 s^-1.'
))
```

At the deeper range, Naples is most similar to the Ionian. At 100 m in the Ionian sea is:
  
```{r}
print(paste(
  tot_phot_irrad$irradiance[tot_phot_irrad$depth==100 & 
                            tot_phot_irrad$sea=="Ionian"],
  'umol m^-2 s^-1.'
))
```
 

 
Multiply the values by 5 (as the binwidth is 5 nm and values are per nm in that range).
 
```{r}
# scalar_irrad_df |> filter(between(wv,400,700)) |>
#   group_by(depth) |>
#   summarise(umol    = signif(5 * sum(irrad),2),
#             umol_sd = signif(5 * sd(irrad),1),
#             quanta  = signif(5 * sum(irrad/1e6)*6.02214076e23,3),
#             ) -> irrad_sum_df
# qplot(data=irrad_sum_df,depth,umol,main='Micromoles of quanta between 400 nm and 700 nm at increasing depth (m)') + theme_classic() #+ ylim(0,.2)
```
 

## Converting Spectral Irradiance to Lux
  
```{r}
irrad2lux <- read.csv('spectral_irradiance_2_lux.csv')
head(irrad2lux)
```
  
Planck's constant * speed of light / wavelength gives the energy of a photon in joules. This times the number of quanta per second gives the Wattage (1 W = 1 J / s). 
 
```{r}
h <- 6.62607015e-34 # Planck's constant
c <- 299792458e9    # speed of light in nm

water_cols |> #select(-irrad_SD) |>
  #mutate(quanta_in_5nm_bin = quanta * 5) |>
  mutate(W_in_5nm_bin = quanta * 5 * h * (c/wv) ) -> water_cols
```
 
Add new 10 nm bins.

So far, this method leaves in the peripheral values...
 
```{r}
bin10nm_fun <- function(wv) {
  bins_10nm <- seq(from=375,to=875,by=10)
  wv10nm <- bins_10nm[which.min(abs(bins_10nm - wv))]-5
  return(wv10nm)
}

water_cols |> 
  mutate(wv10nm = purrr::map_dbl(wv,bin10nm_fun)) -> water_cols
head(water_cols)
```
 
Sum the power into the 10 nm bins. The values were previously multiplied by 5 to reverse the averaging per nm and the totals per 5 nm bin were added to get totals per 10 nm bin. These are below divided by 10 to get per nm averages, i.e. spectral irradiance. 
 
```{r}
water_cols |> group_by(wv10nm,depth,sea) |> 
  summarise(W_in_10nm_bin = sum(W_in_5nm_bin),
            W_in_10nm_per_nm = W_in_10nm_bin/10,
            #quanta_in_10nm_bin = sum(quanta_in_5nm_bin)
            ) -> Watts_10nm_bin
head(Watts_10nm_bin)
```

Taken from Sönke's book

> You can convert an irradiance spectrum E($\lambda$) given in W/m2/nm to lux using:

$$ lux = 637 \sum_{\lambda=380}^{\lambda=780} E(\lambda) * \overline{y} (\lambda) * \Delta \lambda $$

where $\overline{y} (\lambda)$ is the photopic (light-adapted) luminosity curve for humans (given in the table below). In other words, to calculate lux, you multiply your irradiance spectrum (first binned into 10 nm intervals) by the luminosity curve and by $\Delta \lambda$ (which is **10 nm** in this case) and then add up all the products. Then multiply this final sum by 673. This is most easily done in a spreadsheet program. Remember that your irradiance must be in watts, since that’s how the luminosity curve was created. The values of ($\lambda$) below 400 nm and above 700 nm are small, so don’t worry if you don’t have irradiance data for this range. The table is in *'spectral_irradiance_2_lux.csv'*.


The human photometric conversions are in 10 nm bins, whereas the Neapolitan values are in 5 nm bins. Therefore, average the Naples values into 10 nm bins, get as Watts (all above) and calculate. 

Get only relevant depths and wvs

**NEED TO REPLICATE THE Y BAR LAMBDA SET FOR EACH CONDITION OF DEPTH AND LOCATION**
 
```{r}
Watts_10nm_bin |> group_by(sea) |>
  filter(depth == 0, between(wv10nm,380,780)) -> Watts_10nm_0m
Watts_10nm_0m$y_bar_lambda <- irrad2lux$y_bar_lambda
head(Watts_10nm_0m)
```

  
 
```{r}
Watts_10nm_0m |> group_by(sea) |>
  mutate(lux = 637*(W_in_10nm_per_nm)*(y_bar_lambda*10),
        # klx = lux / 1000
         ) -> Watts_10nm_0m
head(Watts_10nm_0m)
```
 
 
 
```{r}
sum(Watts_10nm_0m$klx) 
```
Need to dot this by group. 
This value is slightly below what would be expected from direct sunlight. 
 
 




